### 2024-07-01 article/images/benchmark.tex

\begin{tikzpicture}
    \begin{groupplot}[group style={group size= 2 by 1},height=6cm,width=.45\textwidth]
        \nextgroupplot[
            legend columns=-1,
            legend to name=usageLegend,
            ymin=1,
            ymax=1000,
            ymode=log,
            log basis y=10,
            ylabel={runtime},
            xlabel={vector size},
            y SI prefix=milli,
            y unit=s,
            xmin=27,
            xticklabel={$2^{\pgfmathparse{\tick}\pgfmathprintnumber{\pgfmathresult}}$},
            enlarge x limits=auto,
            grid=both,
            tick align=outside,
            tickpos=left,
            title=Intel,
        ]
        \addplot [
            mark=+,
            error bars/.cd,
                y dir=both,
                y explicit,
        ] table [x=bits,y=mean,y error=error] {benchmarks/framework.dat};
        \addlegendentry{Runtime}

        \addplot [
            mark=+,
            color=red,
            error bars/.cd,
                y dir=both,
                y explicit,
        ] table [x=bits,y=commandMean,y error=commandError] {benchmarks/framework.dat};
        \addlegendentry{Command Execution}

        \addplot [
            mark=+,
            color=blue,
            error bars/.cd,
                y dir=both,
                y explicit,
        ] table [x=bits,y=precomMean,y error=precomError] {benchmarks/framework.dat};
        \addlegendentry{Precomputation}

        \coordinate (topleft) at (rel axis cs:0,1);% coordinate at top of the first plot

        \nextgroupplot[
            ymode=log,
            ymin=1,
            ymax=1000,
            log basis y=10,
            xlabel={vector size},
            xmin=27,
            yticklabel=\empty,
            xticklabel={$2^{\pgfmathparse{\tick}\pgfmathprintnumber{\pgfmathresult}}$},
            enlarge x limits=auto,
            grid=both,
            tick align=outside,
            tickpos=left,
            title=AMD,
        ]

        \addplot [
                mark=+,
                color=black,
                error bars/.cd,
                    y dir=both,
                    y explicit,
            ] table [x=bits,y=mean,y error=error] {benchmarks/hetzner.dat};

            \addplot [
                mark=+,
                color=red,
                error bars/.cd,
                    y dir=both,
                    y explicit,
            ] table [x=bits,y=commandMean,y error=commandError] {benchmarks/hetzner.dat};

            \addplot [
                mark=+,
                color=blue,
                error bars/.cd,
                    y dir=both,
                    y explicit,
            ] table [x=bits,y=precomMean,y error=precomError] {benchmarks/hetzner.dat};

        \coordinate (topright) at (rel axis cs:1,1);% coordinate at top of the second plot

    \end{groupplot}
    \path (topleft)--(topright) coordinate[midway] (group center);
    \node[align=center,above,yshift=.8cm] at(group center) {\pgfplotslegendfromname{usageLegend}};
\end{tikzpicture}